---
title: "Assignment_1"
author: "Stephenson_William"
date: '2023-03-09'
output: 'pdf_document'
---

<h1> Assignment 1 

For the first assignment in Organization network analysis, I analyzed my 
LinkedIn connections to see how they realte with one another. Throughout this 
I upload the document, identify what companies have teh most connections, and 
then define the nodes and edges for this project.

Finally I explore, with varying degrees of success, how to graph these 
relationships.


```{r}

library(igraph)
library(tidygraph)
library(tidyverse)
library(readr)
library(dplyr)
library(ggraph)

Connections <- read_csv("Connections.csv")
View(Connections)


Connections %>% 
  count(Company, name = "Total_Count")

```

```{r}
Connections$Label <- paste0(Connections$`First Name`, " ", substr(Connections$`Last Name`, 1, 1))

```



```{r}
# Create nodes data frame
nodes <- data.frame(name = unique(Connections$Label))

# Create edges data frame
edges <- Connections %>%
  inner_join(Connections, by = "Company") %>%
  filter(Label.x != Label.y) %>%
  select(from = Label.x, to = Label.y)
```


```{r}
# Create graph object
g <- graph_from_data_frame(edges, vertices = nodes)

# Optional: plot the graph
plot(g, vertex.label = V(g)$name)
```

```{r}
# Create tbl_graph object
g2 <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

# Plot the graph
g2 %>%
   ggraph(layout = "kk") +
   geom_edge_arc(colour= "grey",
                 lineend = "round",
                 strength = .1,
                 alpha = .1) +
   geom_node_text(aes(label = name), 
                  repel = TRUE, 
                  point.padding = unit(0.2, "lines"), 
                  colour="red") +
  theme_graph(background = "white") +
  guides(edge_width = FALSE,
         edge_alpha = FALSE) +
  labs(title = "Will Stephenson LinkedIn Network Analysis",
       subtitle = "How my first degree connections are connected to one another")

```

```{r}
library(visNetwork)
library(dplyr)

# Read in the data from a CSV file
Connections <- read.csv("Connections.csv")
Connections$FullName <- paste(Connections$First.Name, Connections$Last.Name, sep = "_")

# Define the nodes and edges
nodes <- Connections %>%
  select(FullName, Company) %>%
  distinct(FullName, .keep_all = TRUE) %>%
  mutate(id = FullName, label = FullName, title = Company)

edges <- Connections %>%
  filter(!is.na(Company)) %>%
  group_by(Company) %>%
  filter(n() > 1) %>%
  select(FullName, Company) %>%
  group_by(Company) %>%
  mutate(from = FullName, to = Company)


# Create the network graph
visNetwork(nodes, edges, width = "100%") %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
             nodesIdSelection = TRUE, selectedBy = "Company") %>%
  visNodes(color = list(background = "#A0CBE2"),
           font = list(face = "verdana", size = 20)) %>%
  visEdges(arrows = "edges") %>%
  visGroups(groupname = "McMaster University", color = "maroon") %>%    # darkblue for group "A"
  visGroups(groupname = "ArcelorMittal Dofasco", color = "orange")  %>%
  visInteraction(navigationButtons = TRUE)
```

```{r}
library(treemap)

Connections <- Connections %>%
  group_by(Company) %>%
  mutate(Value = n()) %>%
  ungroup()

p <- treemap(Connections,
            index=c("Company","Position"),
            vSize="Value",
            type="index",
            palette = "Set2",
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )        

```

