---
title: "Assignment 1"
author: "Emery Dittmer"
date: '2023-03-09'

output:github_document
  pdf_document: default
  html_document: default
---

# LinkedIn Contact Analysis

This is an analysis of my linkedin network. The data data was extracted from linked in and reflects the connections with the companies that they currently work for. We will set up the data in visualizations.

We ideally want to set up the data in such a way that we are connected to a company (edge) which is connected to all the people we know at that compnay (node).


## Pre-Processing
### Library & Data
Before getting started on visualizing the network lets set up the libraries and data we need.

```{r setup}
library(readr)
library(dplyr)
library(tidygraph)
library(reticulate)
library(stringr)
library(ggplot2)
library(ggraph)
library(tidyverse)
library(ggrepel)
use_condaenv("DataSci")
```

Let's import the data.

```{r Data loading}
#Import as Text
my_data <- read.delim('Connections.csv',sep =",", header = TRUE,skip=3)
#Import as CSV
connections=read.csv('Connections.csv',skip=3)

#limit data for simplicity
#connections=connections[1:50,]
```

### Data pre-processing
we will now reshape the data so that we can get the names of companies, the company counts and add ourselvs to the network.


```{r get networks}
connections=drop_na(connections)
connections <- subset(connections, Company != "")

#this datframe shows all the connections with a total!
Connections_Detail_with_total <- connections %>%
  filter(!is.na(Company)) %>%
  #not all blanks were caught by the filter. Requires more robust approach
  filter(length(Company)>2) %>%
  group_by(Company) %>%
  summarise(weight=n()) %>%
  bind_rows(summarise(., across(where(is.numeric), sum),
                         across(where(is.character), ~'Total')))

Connections_Detail <- connections %>%
  filter(!is.na(Company)) %>%
  #not all blanks were caught by the filter. Requires more robust approach
  filter(length(Company)>2) %>%
  group_by(Company) %>%
  summarise(weight=n())

#get the list of names and how the connections fit
Connections_Detail=merge(Connections_Detail,connections,by='Company',all=TRUE)

#add First and last
Connections_Detail <- Connections_Detail %>%
  mutate(peoplenames = paste(First.Name,substring(Last.Name,1,1)))

#set column order
col_order=c("peoplenames","weight","Company","First.Name","Last.Name","Email.Address","Position","Connected.On")

#reorder columns
Connections_Detail <- Connections_Detail[, col_order]

#add to and from fields
Connections_Detail <- Connections_Detail %>%
  mutate(from = Company, 
         to = peoplenames) %>%
  select(from, to, weight)

#get company names for later
companies <- unique(Connections_Detail$from)

#Add yourself to the network
##create a dataframe of yourself
You_network=data.frame(from=rep("you",length(unique(Connections_Detail$from))),to=unique(Connections_Detail$from),weight=rep(1,length(unique(Connections_Detail$from))))

#adding to the concctiond
Connections_Detail=rbind(Connections_Detail,You_network)

```


##Graph Table
### Set data into graph table
In order to visualize the network we need to set the data into a special type of table. The graph table determines relationships between nodes. These can be to and from relationships in directed netwroks.
```{r graph table}
graph_connections <- as_tbl_graph(Connections_Detail)
```

We will need to modify the table and add a title field to make sure that it is accepted by the code that vislizes the data.

```{r nework plotting 2 }
graph_connections <- graph_connections %>%
  activate(nodes) %>%
  mutate(
    title = str_to_title(name),
    label = str_replace_all(title, " ", "\n")
    )

graph_connections
```
## Plotting
### Using GGraph

First we will use ggraph to map some of the connections

```{r network plotting 4}
thm <- theme_minimal() +
  theme(
    legend.position = "none",
     axis.title = element_blank(),
     axis.text = element_blank(),
     panel.grid = element_blank(),
     panel.grid.major = element_blank(),
  ) 

theme_set(thm)
graph_connections %>%
  ggraph(layout = "kk") +
    geom_node_point() +
    geom_edge_diagonal() +
  geom_node_text(aes(label = label, alpha=0.1))
  
```
The cloud is complex, let's try some other methods to clear it up.

```{r netwrok plotting 5}
theme_set(thm)
graph_connections %>%
  ggraph(layout = "kk") +
    geom_node_point() +
    geom_edge_diagonal() +
  geom_node_text(aes(label = label, alpha=0.1))
```


```{r netwrok plotting 6}
lapply(c('stress', 'fr', 'lgl', 'graphopt','kk'), function(layout) {
  graph_connections %>%
    ggraph(layout = layout) +
    geom_edge_fan(width = .2, color = 'lightblue') + 
    geom_node_text(aes(label = label, color = name,alpha=0.1), size = 2) + 
    coord_fixed()+
    scale_fill_brewer()
})
```
```{r netwrok plotting 7}
lapply(c('stress', 'fr', 'lgl', 'graphopt','kk'), function(layout) {
  graph_connections %>%
    ggraph(layout = layout) +
    geom_edge_fan(width = .2, color = 'lightblue') + 
    coord_fixed()+
    scale_fill_brewer()
})
```


```{r netwrok plotting 8}
graph_connections %>%
  ggraph(layout = "stress") +
    geom_node_text(aes(label = label, color = name), size = 2) +
    geom_edge_diagonal(color = "gray", alpha = 0.4) 
```
```{r netwrok plotting 9}
graph_connections %>%
  ggraph(layout = "stress") +
    geom_edge_diagonal(color = "gray", alpha = 0.4) 
```

```{r netwrok plotting 10}
ggraph(graph_connections, 'circlepack') + 
  geom_edge_link() + 
  geom_node_point(aes(colour = depth)) +
  coord_fixed()
```
We can also use a hierarchical or tree typ view

```{r netwrok plotting 11}
ggraph(graph_connections, 'tree') + 
  geom_edge_diagonal()+
  geom_node_text(aes(label = label,alpha=0.2, color='blue'),size=3)
```

```{r netwrok plotting 12}
ggraph(graph_connections, 'dendrogram', circular = TRUE ) + 
  geom_edge_elbow() + 
  #coord_fixed()+
  #geom_node_text(aes(label = label , color = name), size = 2)+
  geom_node_text(aes(label = label , color = ifelse(name %in% companies, "red","black"), size = 2))
```

Attempted a branching or unrooted tree but it was not effective
```{r netwrok plotting 13}
ggraph(graph_connections, 'unrooted') + 
  geom_edge_link()+
  geom_node_text(aes(label = label), size = 2)
```


### Visnetwork

```{r}
require(visNetwork, quietly = TRUE)
nodes <- data.frame(Connections_Detail$from)
edges <- data.frame(from= Connections_Detail$from ,to=Connections_Detail$to)
visNetwork(nodes, edges, width = "100%")
```


```{r}
library(rpart)
require(visNetwork, quietly = TRUE)
#install.packages('sparkline')
library(sparkline)

# Complex tree
data("solder")
res <- rpart(Opening~., data = solder, control = rpart.control(cp = 0.00005))
visTree(res, height = "800px", nodesPopSize = TRUE, minNodeSize = 10, maxNodeSize = 30)


res <- rpart(name~., data = Conection_Detail, control = rpart.control(cp = 0.005))
visTree(res, height = "800px", nodesPopSize = TRUE, minNodeSize = 10, maxNodeSize = 30)


```






### Tidy Graph
We will now use a package called tidy graphy to visualize the netwrok.
**Credit to:http://users.dimi.uniud.it/~massimo.franceschet/ns/syllabus/make/tidygraph/tidygraph.html**

```{r tidy graph 1}
# graph analysis and visualziation
library(tidygraph)
library(ggraph)
library(igraph)

# tidy data analysis and visualziation
library(readr)
library(dplyr)

```

```{r tidy graph 2}
graph <- as_tbl_graph(Connections_Detail)

# plot using ggraph
ggraph(graph, layout = 'kk') + 
    geom_edge_fan(aes(alpha = after_stat(index)), show.legend = FALSE) + 
    geom_node_point() + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')

```


Visualizing the data is tricky but we can do some data manipulations to show more clearly the network.

```{r graph nodes connections}
# edge size shows frequency of co-occurrence
graph %>%
   ggraph(layout = "fr") +
   geom_edge_arc(colour= "gray50",
                  lineend = "round",
                 strength = .1) +
   geom_node_point() +
   geom_node_text(aes(label = name), 
                  repel = TRUE, 
                  point.padding = unit(0.2, "lines"), 
                  colour="gray10") +
  scale_edge_width(range = c(0, 2.5)) +
  scale_edge_alpha(range = c(0, .3)) +
  theme_graph(background = "white") +
  guides(edge_width = FALSE,
         edge_alpha = FALSE)
```
Unforuntely the sizing functions are not currently working but would help determine the strength of network ties.

<!-- Lets try with the name sizes -->

<!-- ```{r graph nodes connections 2} -->
<!-- # edge size shows frequency of co-occurrence -->
<!-- graph %>% -->
<!--    ggraph(layout = "fr") + -->
<!--    geom_edge_arc(colour= "gray50", -->
<!--                   lineend = "round", -->
<!--                  strength = .1) + -->
<!--    geom_node_point(size=(s)*2) + -->
<!--    geom_node_text(aes(label = name),  -->
<!--                   repel = TRUE,  -->
<!--                   point.padding = unit(0.2, "lines"),  -->
<!--                   colour="gray10") + -->
<!--   scale_edge_width(range = c(0, 2.5)) + -->
<!--   scale_edge_alpha(range = c(0, .3)) + -->
<!--   theme_graph(background = "white") + -->
<!--   guides(edge_width = FALSE, -->
<!--          edge_alpha = FALSE) -->
<!-- ``` -->


## Vsiualization II
### Visulize the results in python's plotly
Let's try some python pacakges to enhance what we have

```{python visulization}
import pandas as pd
import numpy as np
from pyvis.network import Network
import plotly.express as px
import networkx as nx
import matplotlib.pyplot as plt
df = pd.read_csv('Connections.csv',skiprows=3)
df['Full Name']= df.apply(
    lambda x: str(x['First Name']) +" "+ str(x['Last Name']) ,
    axis=1)
    #lambda x: str(x['First Name']) +" "+ str(x['Last Name'])[0] , axis=1)
df=df.dropna(subset=['Company', 'Position'])
df=df.groupby("Company").filter(lambda x: len(x) > 1)

df['Count_Company'] = df.groupby('Company')['Company'].transform('size')
df['Count_Company']=df['Count_Company'].astype(int)

df['Position_Count'] = df.groupby('Position')['Position'].transform('size')
df['Position_Count']=df['Position_Count'].astype(int)
df.head()
px.treemap(df, path=['Company', 'Position'],
           width=1000,
           height=1000,
           color='Count_Company',
           color_continuous_scale='YlGnBu',
           color_continuous_midpoint=np.average(df['Count_Company']))
           
px.treemap(df, path=['Position','Company'],
           width=1000,
           height=1000
           )
```


```{python viz 3}

```


```{python vis 4}
net = Network(height="750px", width="100%", bgcolor="#222222", font_color="white",notebook=True,filter_menu=True)

# set the physics layout of the network
net.barnes_hut()

sources = df['Company']
targets = df['Full Name']
weights = df['Count_Company']

edge_data = zip(sources, targets, weights)

for e in edge_data:
                src = e[0]
                dst = e[1]
                w = e[2]
                net.add_node(src, src, title=src)
                net.add_node(dst, dst, title=dst)
                net.add_edge(src, dst, value=w)

neighbor_map = net.get_adj_list()

# add neighbor data to node hover data
for node in net.nodes:
                node["title"] += " Neighbors:<br>" + "<br>".join(neighbor_map[node["id"]])
                node["value"] = len(neighbor_map[node["id"]])

net.show("network.html")
```


